Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-05-20 LaTeX/response.tex

\noindent\subsubsection{}
L195: define the epsilon character, which has not appeared before, or specify the size of the selection parameters relative to recombination, since this is the key point of contrast (e.g., s $<<$ r).

\noindent\paragraph{Response}
\textcolor{blue}{We have replaced this instance of $\epsilon$ with the more familiar $s<<r$ style assumption (L215-216). 
When we introduce $\epsilon$ in the weak selection section we now verbally explain that it is just a number that is much less than one (L310-312).}

@@

\noindent\subsubsection{}
Fig. 2, panel C. You might consider using log scale on the y-axis of the inset, since the blue and green curves do not really show.

\noindent\paragraph{Response}
\textcolor{blue}{This is now Figure 3, panel C. Here the blue and green curves should not really show as they are not expected to increase from a very small initial frequency. We have kept the original scale but now state in the legend that the blue and green loci do not invade.}

@@

\section{Reviewer 3}

This work (henceforth [SOO]) explores the role of both diploid and haploid selection on the evolution of new sex-chromosomes and or transitions between sex-determination systems. For that purpose, the authors develop a population genetics model of three loci with two alleles in each locus. The first locus X is the ancestral sex-determining locus, the second locus A is the locus under diploid and haploid selection. The third locus M is the novel sex-determining locus.

\noindent\subsection{}
I find the article is not clearly written and could benefit from improving the presentation of the main results. 
I also find the article claims that: (a) the modeling work is novel and (b) the findings are unexpected to be only partially justified.
The works of Kozielska 2014 and Ubeda 2015 model the combined effects of both haploid and diploid selection on the evolution of new sex-chromosomes. The work of [SOO] is thus not novel in considering both haploid and diploid selection on the evolution of new sex-chromosomes. It does generalize previous work by considering sex-specific viability selection and get some nice analytical results.

\noindent\paragraph{Response}
%\textcolor{red}{The abstract and summary still seem to be a little over-selling...}
\textcolor{blue}{
We have performed an extensive re-arrangement to improve the clarity of our writing, especially regarding the main results (see Responses 2.1-2.3). 
As suggested by the reviewer (here and below), we also now give more prominence to the related results by Kozielska et al.\ 2010 and Ubeda et al.\ 2015 (e.g., L56-71, L198, L367-373, L440-446). 
In particular, we introduce these studies (L56-71) and more precisely indicate which of our model features and results are new and unexpected in light of these previous results.
}

\textcolor{blue}{
While related, these two studies considered particular cases that we do not consider at length. 
For example, Ubeda et al.\ (2015) consider ESD-GSD transitions. 
Nevertheless, some steps in Ubeda et al.\ (2015) involve increased linkage with a haploid selected locus, which is precedent for our Conclusions 3A and 3B (which are for GSD-GSD transitions in which linkage with a haploid selected locus is increased).
We now mention this in the introduction (L58-60) and results (L369-371). %, and Discussion (L432).  
Similarly, Kozielska et al.\ (2010) involves a special case of GSD transitions during which sex-linkage goes from perfect to none, which is an example of our Conclusion 2 and is similar to Conclusion 3C, which are conclusions regarding GSD-GSD transitions in which linkage with a haploid selected locus is decreased.
This is mentioned on in numerous places in the text (L62-67, L301-304, and L367-368). 
We believe that our new presentation makes the novel and unexpected features of our results much more apparent. 
For example, despite these precedents, we find the combination of Conclusions 3A, 3B, and 3C to be a striking reconciliation of the results of Kozielska et al.\ (2010) and Ubeda et al.\ (2015), see Conclusion 4 as well. 
We elaborate further on what makes our results novel and surprising in the responses below.
}

@@

\noindent\subsubsection{}
The findings in [SOO] claimed to be surprising (``Surprisingly, we find that neither force (sex ratio selection nor associations with genes that have sex-specific effects) dominates the spread of new sex-determining systems alone.") ... [but this result is rather similar to those in Ubeda et al 2015].

\noindent\paragraph{Response}
\textcolor{blue}{
Here we were referring to the result that $\lambda_{W'}^{XY} = \lambda_{Y'}^{ZW}$ with haploid selection, even when sex-ratio biases evolve to become either stronger or weaker (e.g., $\lambda_{Y'}^{ZW}$ vs. $\lambda_{W'}^{XY}$ with $R<r$ and male meiotic drive) or when sex-ratio biases are either present or absent (e.g., $\lambda_{W'}^{XY}$ vs. $\lambda_{Y'}^{ZW}$ with $r<1/2$, $R=1/2$, and male meiotic drive).  
As mentioned by the reviewer, Step 2 in Ubeda et al.\ 2015 shows that associations with a haploid selected locus can favour a new sex-determining allele despite causing biased sex ratios. 
Nevertheless, it was difficult for us to intuit whether the presence/absence of ancestral sex-ratio bias would have a negligible or overwhelming effect on the evolution of new sex-determining alleles. 
Because we start from GSD, our model allows ancestral sex-ratio biases to build up due to the presence of haploid selected loci near to the ancestral sex-determining locus. 
There are reasons to suspect that ancestral-sex-ratio biases would strongly influence the spread of new sex-determining systems.
For example, step 3 of Ubeda et al.\ involves the spread of an autosomal meiotic drive suppressor favoured by Fisherian sex-ratio selection and Kozielska et al.\ 2010 show that ancestral sex-ratio bias can favour new sex determining systems that balance the sex ratio.
Furthermore, reviews mentioned on L21 suggest that Fisherian sex-ratio selection is a particularly dominant force driving transitions between sex-determining systems.
We found our result, that beneficial associations and sex ratio selection exert exactly equal forces on the spread of neo-sex-chromosomes (with selection weak relative to recombination), such that there is no difference between the propensity for XY-ZW or ZW-XY transitions, to be surprising.
To clarify our meaning we have rephrased (in Author summary): ``Surprisingly, we find the two forces (selection to equalize the sex-ratio and the benefits of hitchhiking alongside driven alleles that distort the sex ratio) will often be equally strong, and thus neither is sufficient to explain the spread of new sex-determining systems in every case." 
We believe that the re-framing of the Introduction/Results and the emphasis given to specific Conclusions will help to clarify this point throughout the paper (see Responses 2.1-2.3). 
In particular, this result is now highlighted as `Conclusion 4'.
}

@@

\noindent\subsubsection{}
\noindent and unexpected (``Even more unexpectedly, we find that, to spread, new sex-determining alleles do not necessarily have to arise in closer linkage with genes that are differentially selected in males and females"), are rather similar to those in Ubeda et al 2015. This is a problem that repeats itself over the paper

\noindent\paragraph{Response}
\textcolor{blue}{
The Ubeda et al.\ (2015) results do indeed show that differential selection between male and female diploids is not required for new sex-determining alleles to spread; they show that a new genetic sex-determining allele can spread from ancestral ESD when it is linked to a locus that experiences sex-specific meiotic drive. 
In this sentence, we were attempting to stress that new sex determining alleles can spread when they are \textit{less closely linked or even completely unlinked} to loci under differential male/female haploid and/or diploid selection. 
To our knowledge, this has only been shown for the special case of male meiotic drive in an ancestral XY system with $r=0$ and $R=1/2$ with a neo-W, which is driven by sex-ratio selection (Kozielska et al.\ 2010). 
We therefore found the large parameter space allowing invasion with $r<R$ (Table 3), sometimes in the absence of sex-ratio selection (or even in the absence of haploid selection), to be very unexpected.
To improve our communication of this result, we have rephrased (in Author Summary): ``We also find that new sex-determining alleles can spread despite being less closely linked to selected loci as long as initial linkage is tight or haploid selection is present." 
The changes to the presentation of the Introduction, Results, and Discussion should also help to clarify this point further (see Responses 2.1-2.3). 
}

@@

\noindent\subsubsection{}
For example in: Page 2. The authors claim that predominant theories of the evolution of sex-determining systems are two: sexually-antagonistic selection and sex-ratio selection. They ignore other two: inbreeding (Charlesworth 1978) and sex-different drive (Ubeda et al 2015).

\noindent\paragraph{Response}
\textcolor{blue}{
Here, sexually-antagonistic selection was referred to as an example of ``selective differences between the sexes". 
We believe the phrase `selective differences between the sexes' includes sex-differences in drive (Ubeda et al 2015), which occurs during haploid rather than diploid selection, as indicated in Fig 1B (see also Response 3.1.6).
The enhanced discussion of Ubeda et al.\ (2015) immediately below this statement should help to clarify this point (L58-62).
\\ 
Charlesworth and Charlesworth 1978 (Am Nat and Heredity) discuss the role of inbreeding (and particularly inbreeding depression) in transitions from hermaphroditism to dioecy. 
As described in the model section and lines 468-472 of the original submission, we have only considered GSD-GSD transitions and GSD-ESD transitions.
For clarity, we have moved this description into the Introduction (L48-49). 
When we draw a parallel between ESD and hermaphroditism (L483-491), we suggest future theory include inbreeding and cite Charlesworth and Charlesworth (1978). 
}

@@

\noindent\subsubsection{}
Page 3 ``Here we use mathematical models to find the conditions under which new sex-determining systems spread when individuals experience selection at both diploid and haploid stages, which allows fitness differences between the sexes and sex ratio biases to occur simultaneously".
This would be the same model as Ubeda et al 2015 adding sex-specific viability.

\noindent\paragraph{Response}
%\textcolor{red}{edit}
\textcolor{blue}{
As we now make clear (see response 3.1), we do not look at the same situation as Ubeda et al 2015 and our contribution is not just to add sex-specific viability. 
In particular, Ubeda et al start from ESD. 
Therefore, associations with an ancestral sex-determining locus cannot build up. 
These lead to some of our most surprising results (e.g., haploid selection favours trans-GSD transitions for a wider variety of genomic locations than diploid selection alone; ESD does not invade despite initial sex-ratio bias).
Further, our model and intuition-giving analytical solutions explore a much wider range of scenarios than the numerical results of Ubeda et al., allowing, for example, sex-specific selection (e.g., sex-antagonism) and various genomic arrangements (any order/linkage of the three loci).
Our model is thus much more general and clarifies just when one should expect to see the numerical results of Ubeda et al.
We have rephrased to make this clear (after discussing Ubeda et al.\ 2015 in the Introduction, see Response 2.1; L72-76): ``Here, we analytically find the conditions under which new GSD or ESD systems spread in ancestral GSD systems with any degree of linkage between the loci involved and arbitrary forms of haploid and diploid selection. 
Doing so, we reconcile and generalize the results of Kozielska et al.\ (2010) and Ubeda et al.\ (2015) by deriving conditions for the spread of new GSD systems that alter linkage with haploid selected loci."
}

@@

\noindent\subsubsection{}
Lines 68-69: ``If we assume that haploid selection at any particular locus predominantly occurs in one sex (e.g., meiotic drive during spermatogenesis), then such loci experience a form of sex-specific selection."

This is a stretch and creates confusion. I consider selection acting on individuals with different sex-specific selection, now when selection is acting on gametes (which are neither males or females) males or females are just different environments in which selection may take place. Neither the dynamics nor the constraints are the same.

\noindent\paragraph{Response}
\textcolor{blue}{In this study, we are referring to organisms, such as animals and angiosperms, in which the type of gamete depends on the sex of the diploid that produced it. In this respect, one could consider gametic selection among sperm, for example, to be a form of male-specific selection. 
However, as noted by the reviewer here, the dynamics of selection during gametic selection cannot be captured by a diploid selection coefficient, which is why we specifically model selection upon haploid genotypes.
In our revision, a roughly corresponding statement now reads (L50-51) ``On one hand, haploid selection is typically sex-limited in that it usually occurs among gametes produced by one sex only [38-41].''.
We have also expanded our explanation of the mating system and the way we model haploid selection (see Response 3.2 below).
}

@@

\noindent\subsubsection{}
Lines 70-71: ``In this respect, we might expect that haploid selection would affect transitions between sex-determining systems in a similar manner to sex-specific diploid selection (as explored in [31,32])"

Following up my previous comment Ubeda 2015 explores this and shows that both forms of selection are different. For example sex-antagonistic selection is needed for the invasion of a new sex-determining gene with diploid selection while antagonistic drive in each of the sexes is not needed for the invasion of a new sex-determining gene with haploid selection.

\noindent\paragraph{Response}
%\textcolor{red}{edit}
\textcolor{blue}{In the absence of sexually-antagonistic diploid selection it is true that antagonistic drive is not needed, but \textit{differences} in haploid selection between the sexes are needed (e.g., drive in only one sex, with variations maintained by ploidally-antagonistic selection). 
If drive was the same in both sexes, $\alpha^{m}=\alpha^{f}$ (and all other male/female selection terms also equal), we show that the new sex-determining allele would not invade (e.g., equation 2 and 3). 
In this sense haploid selection is similar to sex-specific diploid selection, and together haploid and diploid selection determine the differences in male vs.\ female fitness, which can drive sex-determination transitions (see Response 3.1.4).\\
This sentence meant to introduce the `sex-specific' aspect of haploid selection in the way Ubeda et al.\ 2015 considers it; associations with alleles favoured in one sex (or the gametes produced by that sex) should favour a new sex-determining allele.
We have rephrased to make this clear (L51-53): ``Therefore, one might expect new sex-determining systems to benefit from close linkage with haploid selected loci, as found for loci that experience diploid-sex-differences in selection [35-37]."
We now follow this sentence up with a paragraph that clarifies its meaning by comparing with the results of Kozielska et al.\ 2010 and Ubeda et al.\ 2015 (L56-71; see also Response 2.1).}

@@

\noindent\subsubsection{}
Lines 77-78: ``It is not immediately clear how the spread of new sex-determining systems would be influenced by the combination of sex ratio biases and associations with haploid selected allele"

This is indeed the focus of Ubeda et al 2015.

\noindent\paragraph{Response}
%\textcolor{red}{edit}
\textcolor{blue}{While Ubeda et al.\ (2015) did focus on sex ratio biases and haploid selected alleles, the system started with ESD, where there are no associations (see Response 3.1.4).
Further, the restricted scope of that numerical study prevents us from generalizing beyond its special case (see Response 3.1.4).
We now present a refurbished argument for why it is difficult to intuit the implications of haploid selection while summarizing the results of Ubeda et al.\ and Kozielska et al.\ 2010 in the Introduction (see Response 2.1).
}

\noindent\subsubsection{}
I will not provide further examples, but it is surprising that the work of Ubeda et al 2015, which is closely related to the one presented here, is cited in page 12 supporting a statement only tangential to the work of Ubeda et al 2015.

\noindent\paragraph{Response}
%\textcolor{red}{edit}
\textcolor{blue}{
Ubeda et al 2015 was in fact cited in the main text (line 396) of the original submission, where we mentioned that it showed that sex-ratio biases could become more severe during a transition.
However, we thank the reviewer for emphasizing the similarities between that study and our own.
We now discuss Ubeda et al.\ 2015 in the Introduction, Results, and Discussion (see Responses above), which we believe clarifies and strengthens our arguments while giving credit to this earlier work.}

@@

\noindent\subsection{}
In terms of modeling the distinction between meiotic drive and gametic competition is unnecessary and only complicates notation. Furthermore, from biological perspective the division seems blurry with molecular biologists considering that meiotic drive only happens in females and never in males and what is often called meiotic drive in males is gamete competition. I would suggest using a single parameter measuring transmission distortion.

\noindent\paragraph{Response}
\textcolor{blue}{
We have chosen not to adjust our models to include a single parameter that measures transmission distortion. 
Given our definition of meiotic drive and gametic competition (L110-120), the two processes are different.
In particular, meiotic drive can only have an effect in heterozygotes.
For example, this means that cases with `sperm killers' and then ``gametic competition" amongst the gametes produced by a single individual are appropriately modelled by our `meiotic drive' parameter. 
For the same reason, ``gametic competition'' of any type in a monogamous mating system corresponds to `meiotic drive' in our model (L116-119). 
On the other hand, in our model, ``gametic competition'' is competition amongst all the gametes produced by one sex (e.g., pollen from many male plants).
Thus, for example, allele frequency changes can occur by gametic competition even if all individuals were homozygous (as long as there is a polymorphism).
We have extended our description of how we use the terms `gametic competition' and `meiotic drive' and how they are modelled in the model section (L110-120), which helps to clarify why they are different and under what circumstances each applies. 
Unfortunately, beyond defining our terms, it was not straightforward to resolve the different usages of the terms by different communities.
}
